---
title: 'A_microphysiological_model_of_human_trophoblast_invasion_during_early_embryo_implantation'
author: Geremy Clair
date: February 16, 2021
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme : cerulean
    code_folding: hide
---

```{r knit_options, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=6, echo = TRUE, results = TRUE, warning = FALSE, message=FALSE) 
options(repos=structure(c(CRAN="https://cloud.r-project.org/")))
setwd("~/Documents/GitHub/Implantation_on_a_chip_proteomics/")
```

# Required libraries

The required libraries are loaded 
- RomicsProcessor written by Geremy Clair (2021) is used to perform trackable transformation and statistics to the dataset 
```{r load_packages}
library("RomicsProcessor")
library("DT") #for the rendering of the enrichment tables 
library("proteinminion") #this package was created by Geremy Clair (2021) to download UniProt protein details
```

# Fasta and protein ontologies download using 'Protein Mini-On'

Using the package 'Protein Mini-on' (Geremy Clair 2021, in prep.), The fasta file was downloaded from Unipro for the human and bovine proteome on the Jun 15th, 2021
```{r download_fasta}
if(!file.exists("./03 - Output files/Uniprot_Bos_taurus_proteome_UP000009136_2020_06_15.fasta")){
  download_UniProtFasta(proteomeID = "UP000009136",reviewed = F,export = TRUE, file="./03 - Output files/Uniprot_Bos_taurus_proteome_UP000009136_2020_06_15.fasta")
}

if(!file.exists("./03 - Output files/Uniprot_Homo_sapiens_proteome_UP000005640_2020_06_15.fasta")){
    download_UniProtFasta(proteomeID = "UP000005640",reviewed = F,export = TRUE, file="./03 - Output files/Uniprot_Homo_sapiens_proteome_UP000005640_2020_06_15.fasta")
}
```

# MaxQuant import

The iBAQ data contained in the protein table was loaded, the corresponding metadata was loaded
```{r data_metadata_load}
data<-data.frame(extractMaxQuant("./01 - Source files/proteinGroups.txt",quantification_type = "iBAQ",cont.rm = T,site.rm = T,rev.rm = T))
IDsdetails<-extractMaxQuantIDs("./01 - Source files/proteinGroups.txt",cont.rm = T,site.rm = T,rev.rm = T)
IDsdetails<-cbind(UniProt_Name=sub(".*\\|","",IDsdetails$protein.ids), IDsdetails)
colnames(data)<- sub("iBAQ.","",colnames(data))
metadata<- read.csv(file = "./01 - Source files/metadata.csv")
colnames(metadata)<-tolower(colnames(metadata))
write.csv(extractMaxQuantIDs("./01 - Source files/proteinGroups.txt",cont.rm = T,site.rm = T,rev.rm = T),"./03 - Output files/MaxQuantIDS.csv")
```

# Romics_object creation

The data and metadata were placed in an romics_object, the sample names were retrieved from the metadata, the condition  will be use for the coloring of the Figure and statistics
```{r create_romics}
romics_proteins<- romicsCreateObject(data,metadata,main_factor = "Condition")
romics_proteins<- romicsSampleNameFromFactor(romics_proteins,factor = "sample_names")
```


# Full data analysis
## Data cleaning and normalization

The missingness was evaluated for each channel/sample
```{r missingness_eval_full}
romics_proteins<- romicsZeroToMissing(romics_proteins)
romicsPlotMissing(romics_proteins)
```


The proteins to be conserved for quantification were selected to contain at least 70% of complete values (3/4 samples) for a given condition, the overall missingness was evaluated after filtering.
```{r missing_filter_full}
romics_proteins<-romicsFilterMissing(romics_proteins,percentage_completeness = 75)
print(paste0(nrow(romics_proteins$data),"/", nrow(romics_proteins$original_data)," proteins remained after filtering", " (",round(nrow(romics_proteins$data)/nrow(romics_proteins$original_data)*100,2),"%)."))
romicsPlotMissing(romics_proteins)
```

The data was log2 transformed, the distriution boxplot were then plotted
```{r log2_full}
romics_proteins<-log2transform(romics_proteins)
distribBoxplot(romics_proteins)
```

As the same quantity of protein was labelled for each sample, the expectation is that the distribution of the protein abundance is centered, therefore a median centering was performed prior to plot again the distribution boxplots.
```{r median_norm_full}
romics_proteins<-medianCenterSample(romics_proteins)
distribBoxplot(romics_proteins)
```

The grouping of the samples by is checked by hierarchical clustering
```{r hclust_no_imput_full}
romicsHclust(romics_proteins)
```

## Data imputation

For some of the subsequent statistics imputations are required, we performed an imputation by assuming that the "non-detected" proteins were either low abundance or missing using the method developped by Tyranova et al. (PMID: 27348712). The gray distribution is the data distribution, the yellow distribution is the one for the random values used for imputation.
```{r imput_missing_full}
imputeMissingEval(romics_proteins,nb_stdev = 2,width_stdev = 0.5, bin=1)
romics_proteins<-imputeMissing(romics_proteins,nb_stdev = 2,width_stdev = 0.5)
```

The PCA grouping was checked again after imputation
```{r grouping_after_imputation_full}
indPCAplot(romics_proteins, plotType = "percentage")
indPCAplot(romics_proteins, plotType = "individual",Xcomp=1,Ycomp =2)
indPCAplot(romics_proteins,  plotType = "individual",Xcomp=2,Ycomp =3)
indPCA3D(romics_proteins)

```

## Statistics

The means and stdev are calculated for each group
```{r Means_stdevs_full}
romics_proteins<-romicsMean(romics_proteins)
romics_proteins<-romicsSd(romics_proteins)
```

Some general statistics are performed (ANOVA, T.tests).
```{r statistics_full}
romics_proteins<-romicsANOVA(romics_proteins)
romics_proteins<-romicsTtest(romics_proteins,var.equal = T)
print(paste0(sum(romics_proteins$statistics$ANOVA_p<0.05), " proteins had an ANOVA p<0.05."))
```

A heatmap depicting the proteins passing an ANOVA p<0.05 is plotted, the clusters obtained were saved in the statistics.
```{r Heatmap_full}
romicsHeatmap(romics_proteins,variable_hclust_number = 4,ANOVA_filter = "p", p=0.05,sample_hclust = F)
romics_proteins<-romicsVariableHclust(romics_proteins,clusters = 4,ANOVA_filter = "p",p= 0.05,plot = F)
romics_proteins<-romicsZscores(romics_proteins)
```

The data was exported 
```{r export_results}
results<-romicsExportData(romics_proteins,statistics = T,missing_data = T)
write.csv(results, "./03 - Output files/implantation_proteomics_complete_results.csv")
```


